\documentclass[11pt]{article}

\usepackage[margin=1.15in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{microtype}
\usepackage{booktabs}

\hypersetup{
  colorlinks=true,
  linkcolor=blue!45!black,
  citecolor=blue!45!black,
  urlcolor=blue!45!black
}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Lean}{Lean~4}
\newcommand{\mathlib}{mathlib}
\newcommand{\Faces}{\mathcal F}
\newcommand{\Pairs}{\mathcal P}
\newcommand{\rank}{\operatorname{rank}}
\newcommand{\unrank}{\operatorname{unrank}}
\newcommand{\dotp}{\mathbin{\cdot}}

\title{A Formal Proof that the Cuboctahedron Has No Nonsingular Omnihedral Periodic Billiard Orbit}
\author{Zach Conn}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
We give a formal proof strategy for the theorem that the cuboctahedron admits
no nonsingular periodic billiard orbit touching each face exactly once in one
period.  The geometric proof is short in outline: unfold the billiard path to a
straight line, enumerate the possible face orders, and show that each order is
incompatible with the resulting exact affine return map.  The difficulty is
that the finite enumeration contains nearly one hundred million opposite-face
orders and more than one hundred fifty million signed translation cases.

The Lean development handles this finite problem by exact rational arithmetic
and proof-carrying obstruction certificates.  Non-identity return maps are
eliminated by forced-axis obstructions.  Translation return maps reduce to
strict two-variable linear systems; these are eliminated by GoodDirection
arguments, Farkas certificates, Walsh sign-polynomial subcube bounds, and a
semantic descriptor layer for families of two-source contradictions.  External
programs may propose certificates and families, but the finite enumeration,
certificate soundness, and coverage statements are checked in Lean.

At the time of this draft, the public Lean theorem proves the final billiard
statement conditionally from a complete generated coverage witness.  The
geometry, unfolding bridge, pair-word enumeration, certificate checkers,
GoodDirection theorem, Farkas soundness, Walsh denominator infrastructure,
Boolean GoodDirection equivalence, descriptor-family adapters, and coverage
assembly interfaces are formalized.  The remaining publication step is to
expose full Lean-checked generated coverage over all ranks and sign masks.
\end{abstract}

\tableofcontents

\section{Introduction}

A polyhedral billiard orbit is a piecewise-linear path inside a polyhedron
which reflects specularly from faces.  At a nonsingular impact, the incoming and
outgoing directions make equal angles with the face plane; singular impacts, at
edges or vertices, are excluded.  An orbit is \emph{omnihedral} if, during one
period, it visits every face of the polyhedron exactly once.

This paper concerns the cuboctahedron.  The cuboctahedron has fourteen faces:
six squares and eight equilateral triangles.  It is one of the Archimedean
solids, and more specifically the rectification of the cube and the
octahedron.  Thus it is a natural test case immediately beyond the Platonic
solids: it retains a large symmetry group and regular faces, but it has two
different face types.  The theorem under formalization is negative.

\begin{theorem}[Main theorem]
The cuboctahedron has no nonsingular periodic billiard orbit which touches each
of its fourteen faces exactly once per period.
\end{theorem}

The proof is intended to be a formal proof rather than a numerical search.  The
search space is finite but large: after normalizing the starting face, there
are \(97{,}297{,}200\) possible opposite-face pair words, and
\(157{,}957{,}632\) translation sign assignments in the identity-linear branch.
The central engineering problem is therefore not merely to discover obstructions
but to represent them in a form that \Lean can check within realistic memory
and compilation constraints.

\subsection*{Mathematical context and significance}

The unfolding method used here is classical in billiards: instead of reflecting
the moving particle, one reflects the table, so that the billiard path becomes
a straight line in a reflected development of the domain.  In planar polygonal
billiards this construction leads to the theory of rational billiards and
translation surfaces; in that setting Masur proved the abundance of periodic
directions for rational polygons, and Boshernitzan, Galperin, Krueger, and
Troubetzkoy strengthened the picture by proving density of periodic points for
rational polygons \cite{masur,boshernitzan}.  At the same time, even the
apparently elementary question of whether every triangle has a periodic
billiard orbit remains a major open problem.  Billiards are therefore a place
where simple reflection laws produce unexpectedly delicate global questions.

Polyhedral billiards are harder than planar polygonal billiards in several
ways.  The singular set is not a finite set of vertices but a graph of edges
and vertices, the reflected copies live in three-dimensional space, and the
unfolding does not reduce the problem to the same compact translation-surface
picture.  Recent work on polyhedral billiards, such as the study of periodic
tubes avoiding edge neighborhoods by Cekic, Georgiev, and Mukherjee
\cite{cekic-georgiev-mukherjee}, illustrates that even controlling families of
periodic trajectories away from the nonsmooth boundary is a subtle problem.

The omnihedral condition isolates a particularly rigid class of periodic
orbits.  It asks for a shortest possible global tour of the face set: every
face must be hit, and no face may be repeated in the period.  In regular
simplices, such orbits exist; Bedaride and Rao prove, in all dimensions, a
period \(n+1\) billiard trajectory in the regular \(n\)-simplex that hits each
face once \cite{bedaride-rao}.  For highly symmetric polyhedra, it is therefore
natural to ask whether there are analogous face-covering periodic trajectories.
The cuboctahedron is a compelling next case because it is almost as symmetric
as the cube and octahedron from which it is derived, but its mixture of square
and triangular faces introduces new combinatorics.  A negative result says that
large symmetry and regular faces do not, by themselves, force the existence of
an omnihedral billiard orbit.

The formalization is also mathematically significant as a proof style.  The
theorem is not proved by drawing a single clever impossible picture.  Instead,
it is proved by reducing the geometry to a finite exact decision problem and
then making every step of that decision problem proof-carrying.  The result is
a human mathematical argument whose large finite component is checked rather
than trusted.

\subsection*{Why a direct proof is hard}

There is a tempting direct attack.  Choose an order of the fourteen faces,
unfold the cuboctahedron along that order, and ask whether a straight line can
cross the unfolded copies in the prescribed sequence while remaining in the
relative interiors of the faces.  If this fails for every order, the theorem is
proved.

The difficulty is scale.  Even after cyclically starting on \(X_+\) and
forgetting signs inside opposite face-pairs, there are \(97{,}297{,}200\)
valid pair words.  In the translation branch, each identity-linear pair word
has \(64\) sign masks.  A direct proof that writes down and checks one
geometric calculation per signed itinerary would be unreadable on paper and
unbuildable as a formal proof.

There is a second difficulty: most individual cases are not hard for the same
reason.  Some fail because the affine return map has no compatible axis.  Some
fail because the forced direction crosses a face from the wrong side.  Some
fail only after reducing to a two-variable system of strict inequalities.
Among those, many contradictions use the same algebraic shape but with
different coefficients.  A useful proof must therefore expose structure across
families of cases, not merely enumerate cases.

Finally, a computer search is not a proof unless the connection between the
search and the theorem is itself checked.  It is easy for a program to say
``all cases passed''; it is much harder to produce evidence that a proof
assistant can verify without rerunning an enormous untrusted search.  This is
why the formal development emphasizes rank/unrank enumeration, soundness
theorems for certificate checkers, semantic family theorems, interval coverage,
and exact rational arithmetic.

\subsection*{Proof at a glance}

Here is the proof in its human form.  Suppose, for contradiction, that a
nonsingular omnihedral period-14 orbit exists.  Rotate the period so that its
first impact is on the square face \(X_+\).  Unfold the cuboctahedron across
the fourteen hit faces.  The original broken billiard path becomes a straight
line, and the sequence of fourteen reflections becomes a single affine return
map
\[
A(p)=Mp+b.
\]

The started orbit determines a valid opposite-face pair word.  Since every
valid pair word has a Lean-checked rank, it is enough to rule out each rank.
There are two cases.

If \(M\ne I\), the straight line must be compatible with a nontrivial affine
isometry.  This forces an axis, or proves that no such axis exists.  The
non-identity certificate checks this forced situation and shows that the
candidate either fails to exist, starts outside \(X_+\), crosses in the wrong
direction, hits the wrong face, or leaves a required face interior.

If \(M=I\), the return map is a translation.  The starting point is
\((1,y,z)\), and the itinerary conditions are strict linear inequalities in
\(y,z\).  A feasible orbit would give a point satisfying all these
inequalities.  Bad-direction cases contradict a general positivity condition
on internal crossing denominators.  The remaining cases are killed by Farkas
certificates, by Walsh subcube bounds that prove one of those denominators is
nonpositive on an entire family of masks, or by descriptor-level family
theorems that identify the two geometric rows responsible for the
contradiction.

Thus every possible started unfolded itinerary is impossible.  The unfolding
reduction then contradicts the assumed billiard orbit.

\paragraph{Formal status.}
The current public Lean theorem is conditional on complete generated coverage:
\[
\begin{gathered}
\texttt{Cuboctahedron.conditional\_cuboctahedron\_no\_omnihedral} \\
\texttt{(coverage : ExhaustiveGeneratedCoverage) :} \\
\texttt{\quad \textasciitilde{} exists o : BilliardOrbit14,}\\
\texttt{\quad\quad o.Nonsingular /\ o.Periodic /\ o.TouchesEachFaceExactlyOnce.}
\end{gathered}
\]
In words, if Lean is given a complete generated coverage witness, the already
formalized geometry and unfolding bridge prove the billiard theorem.  The
preferred completion path is semantic: generated files should prove compact
impossibility theorems over ranges and families rather than expose large arrays
of raw certificate data.  The latest infrastructure includes a Boolean
GoodDirection bridge and a descriptor-family bridge.  Here ``Boolean'' means a
finite true/false checker proved equivalent to the corresponding mathematical
predicate; it is not an unchecked computation.  The descriptor-family bridge
packages the source rows and row-template identity for a two-source Farkas
contradiction and then erases that package back to the semantic coverage
interface used by the conditional theorem.

\paragraph{Repository.}
The formalization is available at
\[
  \url{https://github.com/zpconn/cuboctahedron}
\]
and this draft refers to repository commit
\[
  \texttt{dff526e891279a0faf6ca83520a47ac4d45981b0}.
\]
The current axiom report for the conditional public theorem is produced by
\[
  \texttt{lake env lean Cuboctahedron/PrintConditionalAxioms.lean}.
\]
When the unconditional generated coverage root is installed, the final
verification target is expected to be
\[
  \texttt{lean Cuboctahedron/PrintAxioms.lean}.
\]

\paragraph{Contribution.}
The contribution is threefold.
\begin{enumerate}[leftmargin=2em]
\item We formalize the exact rational geometry of the cuboctahedron, including
  faces, interiors, reflections, and unfolding into straight-line conditions.
\item We formalize a complete finite enumeration of started omnihedral
  opposite-face pair words by rank and unrank functions checked in Lean.
\item We develop exact certificate checkers and reusable certificate families
  for the two affine-return cases: non-identity linear part and translation.
  The translation branch includes strict two-dimensional Farkas certificates,
  GoodDirection reduction, two-source supports, symmetry transport, Walsh
  sign-polynomial infrastructure for denominator subcube arguments, and a
  descriptor-level coverage language for compact generated families.
\end{enumerate}

\section{The Coordinate Model}

We use the standard coordinate model
\[
P=\{(x,y,z)\in\R^3:
 |x|\le 1,\ |y|\le 1,\ |z|\le 1,\
 \pm x\pm y\pm z\le 2\}.
\]
The square faces are
\[
X_\pm:\ \pm x=1,\qquad
Y_\pm:\ \pm y=1,\qquad
Z_\pm:\ \pm z=1,
\]
and the triangular faces are
\[
T_{\epsilon_1\epsilon_2\epsilon_3}:
  \epsilon_1x+\epsilon_2y+\epsilon_3z=2,
  \qquad \epsilon_i\in\{\pm1\}.
\]

Every face is represented by a rational normal vector \(n\in\Q^3\) and a
rational offset \(c\in\Q\), so that the face plane is \(n\dotp p=c\).  Reflection
in that face is the affine map
\[
r_F(p)=p-2\frac{n\dotp p-c}{n\dotp n}n.
\]
For the cuboctahedron model above, all coefficients in these reflections are
rational.  Thus the formalized reflection maps live in a small exact affine
space over \(\Q\), represented in Lean by rational \(3\times 3\) matrices and
rational vectors.

\begin{remark}
The proof deliberately avoids floating point.  External scripts may profile,
search, or generate candidate data, but the checked objects reduce to exact
integer and rational equalities or inequalities inside Lean.
\end{remark}

\section{Unfolding Billiard Orbits}

The standard unfolding construction replaces a reflected billiard path by a
straight line through reflected copies of the polyhedron.  For a proposed
itinerary
\[
F_0,F_1,\ldots,F_{13},
\]
we compose the corresponding affine reflections and obtain an affine return map
\[
A(p)=Mp+b.
\]
A periodic billiard orbit with this itinerary can exist only if the unfolded
straight line is compatible with \(A\).  The formalization records this
compatibility as \texttt{UnfoldedFeasible}.

Since an omnihedral orbit touches every face once, a period can be cyclically
reindexed to start at a chosen square face.  We choose \(X_+\), called
\texttt{Face.xp} in the Lean development.

\begin{proposition}[Unfolding reduction, formalized]
If a nonsingular omnihedral period-14 billiard orbit exists in the
cuboctahedron, then there is a started face sequence
\[
F_0=X_+,\quad F_1,\ldots,F_{13}
\]
which is omnihedral and admits an unfolded feasible straight-line witness.
\end{proposition}

In Lean this bridge is used in
\texttt{conditional\_cuboctahedron\_no\_omnihedral}.  The theorem first rotates
the orbit to start at \texttt{Face.xp}, then applies
\texttt{billiard\_implies\_unfolded}, and finally invokes the no-started
unfolded-itinerary theorem supplied by generated coverage.

\section{Finite Enumeration of Itineraries}

Opposite faces have the same linear reflection part.  Therefore the proof first
records only which opposite face-pair is hit.  Let the pair set be
\[
\Pairs=\{x,y,z,d_{111},d_{11-},d_{1-1},d_{-11}\},
\]
where \(x,y,z\) denote the three square opposite-pairs and the four \(d\)'s
denote the four triangular opposite-pairs.

After fixing \(F_0=X_+\), a valid pair word has length \(13\) and contains
\[
x\text{ once},\qquad
y\text{ twice},\qquad
z\text{ twice},
\]
and each of the four triangular opposite-pairs twice.  Thus the number of valid
pair words is
\[
\frac{13!}{2^6}=97{,}297{,}200.
\]

The Lean development proves rank/unrank theorems for these pair words.  A rank
is an element of \(\{0,\ldots,N-1\}\), where \(N=97{,}297{,}200\), and
\(\unrank(r)\) is the corresponding valid pair word.  Conversely, every valid
pair word is shown to occur at its rank.  Consequently, a theorem quantified
over \texttt{Fin numPairWords} is a theorem over the complete finite itinerary
space, not over a list trusted from an external search program.

\subsection{Sign masks}

A pair word forgets which face inside each opposite pair is hit.  In the
translation branch, those choices are represented by a six-bit sign mask.  The
initial \(x\)-face has already been fixed as \(X_+\), and the remaining six
opposite-pairs each contribute one binary choice.  Thus every identity-linear
pair word has \(64\) possible translation sign masks.

The proof splits each pair word according to the linear part \(M\) of the
unfolded affine return map:
\[
M\ne I\qquad\text{or}\qquad M=I.
\]
The two cases use different obstructions.

\section{The Non-Identity Case}

Assume \(M\ne I\).  Then a compatible periodic unfolded line is highly
constrained by the affine map \(p\mapsto Mp+b\).  Its direction must be a
nonzero fixed direction of \(M\), and the possible axis of the affine return is
determined by exact linear conditions.  In many cases, even the signed face
sequence and starting point are forced.

The certificate type \texttt{NonIdCert} packages this forced data together with
one of several failure modes.  Typical failures include:
\begin{itemize}[leftmargin=2em]
\item no nonzero fixed direction exists;
\item the required crossing direction has the wrong sign;
\item the forced signed sequence is not omnihedral;
\item the forced axis misses the relative interior of \(X_+\);
\item the candidate hits the wrong face first;
\item an impact lies outside the required face interior.
\end{itemize}
The checker \texttt{checkNonIdCert} verifies the exact data, and the soundness
theorem proves that a passing certificate rules out any real unfolded feasible
orbit realizing the corresponding pair word.

\begin{proposition}[Non-identity obstruction interface]
Given complete generated non-identity coverage, every valid pair word whose
unfolded linear part is not the identity is impossible.
\end{proposition}

The current generated architecture also includes parametric non-identity
families and computable classifiers.  Here a classifier means an exact
decision procedure that assigns a rank to a previously proved family.  It
first removes broad families such as bad-direction cases and bad-pair-balance
cases; the latter are cases in which the forced signed face choices cannot
match the required one-hit-per-face condition.  The decisions are Boolean
checks, meaning finite true/false computations, and their soundness is proved
in Lean.  Residual ranks are then handled by generated interval coverage.

\section{The Translation Case}

Assume \(M=I\).  Then the affine return map is a pure translation,
\[
A(p)=p+b.
\]
The starting point on \(X_+\) has the form
\[
p_0=(1,y,z).
\]
For a fixed pair word and sign mask, the requirements that the unfolded line
cross the prescribed faces in the prescribed order and through the relative
interiors of the copied faces become strict linear inequalities in \(y,z\).

\subsection{Strict linear constraints}

Lean stores a strict two-variable rational linear inequality as a triple
\((a,b,c)\in\Q^3\), interpreted as
\[
ay+bz<c.
\]
The module \texttt{LinearConstraints.lean} contains this language and its
elementary exact arithmetic.  A translation certificate supplies, or
reconstructs from named geometric sources, a finite list of such inequalities.

\subsection{Farkas certificates}

The translation obstruction uses Farkas-style infeasibility certificates.  In
the present two-dimensional strict setting, a Farkas certificate is a finite
list of pairs
\[
(i_j,q_j),
\]
where \(i_j\) indexes an inequality and \(q_j\in\Q\) is a multiplier.  Lean
checks:
\begin{enumerate}[leftmargin=2em]
\item every \(i_j\) is a valid row index;
\item every \(q_j\) is nonnegative;
\item at least one \(q_j\) is positive;
\item after scaling and summing the selected rows, the \(y\)- and
  \(z\)-coefficients are both zero;
\item the summed right-hand side is nonpositive.
\end{enumerate}
If all rows held at some \((y,z)\), then the positive weighted sum would imply
\[
0<c\le 0,
\]
which is impossible.  The theorem
\texttt{checkFarkas\_sound} formalizes this argument.

The development also uses \emph{source-indexed} Farkas certificates.  Instead
of merely naming row numbers, a source-indexed certificate names the geometric
origin of a row, for example an \(X_+\)-start interior condition or an interior
condition at a specified impact.  Lean reconstructs the ordinary row list from
these sources and then invokes the same Farkas checker.

\subsection{GoodDirection}

Before applying full Farkas evidence, the translation branch removes cases
whose crossing directions are already impossible.  For an internal impact, the
formula for the crossing time contains a rational denominator.  A physically
possible translation orbit must have all such internal denominators positive.
The predicate \texttt{GoodDirection} states precisely this positivity
condition.

\begin{proposition}[GoodDirection necessity]
Every feasible translation unfolded orbit satisfies \texttt{GoodDirection}.
\end{proposition}

Consequently, a sign mask that violates \texttt{GoodDirection} needs no
generated Farkas certificate.  It is eliminated by the general theorem.  The
generated translation evidence only needs to cover the surviving
\texttt{GoodDirection} masks.

\subsection{Boolean GoodDirection}

For generated interval proofs it is useful to have a computable version of the
same predicate.  The proof therefore defines
\[
\texttt{goodDirectionAtRankBool}(r,m),
\]
a Boolean value obtained by checking the thirteen internal denominator
inequalities for the rank \(r\) and sign mask \(m\).  A Boolean value is just a
truth value, \texttt{true} or \texttt{false}; the point of this definition is
not to replace the mathematical predicate by computation, but to give Lean a
small finite object that can be used by generated classifiers.

The latest development proves both directions:
\[
\texttt{GoodDirectionAtRank}(r,m)
\quad\Longleftrightarrow\quad
\texttt{goodDirectionAtRankBool}(r,m)=\texttt{true}.
\]
The forward direction was already enough to show that feasible orbits satisfy
the Boolean check.  The new converse is a finite case split over the thirteen
internal impacts.  It does not unfold the rank-specific denominator arithmetic;
it only says that if the finite list of positivity checks is true, then each
of the thirteen positivity inequalities holds.  This equivalence lets a
generated file use whichever surface is smaller: a semantic
\texttt{GoodDirectionAtRank} hypothesis or its Boolean counterpart.

\subsection{Two-source supports}

Many current translation survivors are killed by very small contradictions.  A
two-source Farkas support names two geometric rows and supplies the rational
weights needed to make the \(y,z\)-coefficients cancel.  The helper module
\texttt{TwoSourceFarkas.lean} packages this common proof shape.  A generated
leaf stores a \texttt{TwoSourceFarkasSupport}; Lean reconstructs the two
constraints, checks the Farkas arithmetic, and obtains a semantic theorem that
the case is impossible.

\subsection{Descriptor families}

The newest translation-family layer packages two-source arguments at a more
semantic level.  A \emph{descriptor} is a small record containing three pieces
of data:
\begin{enumerate}[leftmargin=2em]
\item the two indices of the geometric constraint rows in the generated
  constraint-source list;
\item the two-source support naming the geometric origins of those rows;
\item a row template, meaning a reusable algebraic pattern showing how those
  two rows combine into a contradiction.
\end{enumerate}
For example, a descriptor may say that one row comes from a triangular-face
interior condition at the first internal impact, the other from the starting
square \(X_+\), and that the two rows fit a template in which one row has a
fixed positive variable pattern and the other has the opposite fixed pattern.
The exact names of the templates are implementation details; mathematically,
each template is a short rational linear-algebra lemma.

The descriptor predicate has two obligations.  First, the named source indices
must really select the stated geometric rows for the rank and mask under
consideration.  Second, those rows must satisfy the template's algebraic
conditions.  Once both obligations are proved, Lean obtains the same semantic
conclusion as before: the corresponding GoodDirection translation case is
impossible.

This layer is important for publication-quality generated coverage because it
separates three tasks that were previously intertwined.  External generation
may choose a descriptor family.  Lean checks that the descriptor applies to the
particular ranks and masks in the interval.  The hand-written bridge then
converts descriptor coverage into the public theorem
\[
\texttt{AllTranslationGoodCoverageOnRange}(l,h),
\]
which says that every identity-linear GoodDirection survivor in the half-open
rank range \([l,h)\) is killed.  Thus the public proof surface remains a
mathematical coverage statement rather than a listing of case-local
arithmetic.

\section{Walsh Sign Polynomials}

The Walsh infrastructure is a compact way to reason about all \(64\) sign masks
for a fixed pair word without branching over all masks in Lean.  Here ``Walsh''
means an exact sign-polynomial language over six variables.  Each sign bit is
viewed as a value in \(\{1,-1\}\), rather than as a Boolean.

Let
\[
s_y,s_z,s_{111},s_{11-},s_{1-1},s_{-11}\in\{1,-1\}
\]
be the six sign variables.  A Walsh monomial is a product of zero, one, or two
of these variables in the denominator path, and a Walsh polynomial is a
rational linear combination of such monomials.  This is not a numerical
approximation; it is exact rational algebra.

\subsection{Subcubes}

A subcube fixes some sign variables and leaves the others free.  For example,
\[
s_y=1,\qquad s_{111}=-1
\]
with the other four variables free describes \(16\) masks.  A Walsh subcube
obstruction proves that an internal denominator is nonpositive for every mask
in such a subcube.  Since \texttt{GoodDirection} requires that denominator to
be positive, all masks in the subcube are impossible.

The basic Lean structures are:
\begin{center}
\begin{tabular}{ll}
\toprule
Concept & Lean name \\
\midrule
Sign variable & \texttt{SignBit} \\
Boolean subcube & \texttt{MaskSubcube} \\
Walsh monomial & \texttt{WalshMonomial} \\
Walsh polynomial & \texttt{WalshPoly} \\
Subcube upper bound & \texttt{WalshSubcubeUpperBound} \\
Denominator obstruction & \texttt{WalshImpactObstruction} \\
\bottomrule
\end{tabular}
\end{center}

A \texttt{WalshSubcubeUpperBound} contains one rational upper bound per term,
proofs that each term is below its bound on the subcube, and a proof that the
sum of the bounds is nonpositive.  The hand-written theorem then proves that
the whole polynomial is nonpositive on the subcube.

\subsection{Quadratic coefficient records}

The current denominator path uses the compact record \texttt{WalshQuadratic}.
It stores the \(22\) possible coefficients of a degree-at-most-two sign
polynomial:
\[
1,\quad 6\text{ one-bit monomials},\quad 15\text{ two-bit monomials}.
\]
The degree bound is structural.  The copied impact normal is affine in the sign
bits, and the translation vector is affine in the sign bits.  The denominator
is a dot product of these two affine expressions, hence has degree at most two.

Using a fixed coefficient record avoids asking Lean to unfold a generic
polynomial sum for every generated case.  The corresponding upper-bound object
\texttt{WalshQuadraticSubcubeUpperBound} proves nonpositivity slot-by-slot over
the \(22\) coefficients.

\subsection{Symbolic denominator bridge}

The important trust question is why the Walsh expression is the same as the
geometric denominator.  The symbolic bridge avoids a \(64\)-case Boolean split.
The structure \texttt{WalshAffine} represents an affine sign expression, and
\texttt{WalshAffineVec3} represents a vector of such expressions.  The module
\texttt{TranslationWalshVector.lean} mirrors the ordinary exact translation
recurrence, producing a Walsh-affine translation vector.  The module
\texttt{ImpactSubcubeWalshSymbolic.lean} proves that the dot product of a
Walsh-affine impact normal with the Walsh-affine translation vector evaluates
to the claimed \texttt{WalshQuadratic}.  Finally,
\texttt{ImpactSubcubeWalshCompactDenomBridge.lean} connects this dot product
back to \texttt{impactDenomAtRank}.

The resulting trusted path is:
\[
\begin{gathered}
\text{Walsh-affine recurrence for normal and translation vector}\\
\Longrightarrow \text{exact quadratic dot-product coefficients}\\
\Longrightarrow \text{subcube upper bound}\\
\Longrightarrow \text{denominator}\le 0\\
\Longrightarrow \neg\texttt{GoodDirection}.
\end{gathered}
\]
External scripts may propose coefficients, subcubes, and bounds, but Lean
checks the recurrence equalities, coefficient equalities, subcube membership,
and nonpositivity proof.

\section{Symmetry and Semantic Coverage}

The cuboctahedron has \(48\) signed coordinate symmetries when
orientation-reversing symmetries are included.  After fixing the first face to
be \(X_+\), only the subgroup preserving that face remains available.  This
started-face symmetry group has \(8\) elements: it swaps \(y\) and \(z\) and
independently flips their signs.

Symmetry is used only as a compression mechanism.  A symmetry-reduced proof
must still prove, in Lean, that:
\begin{enumerate}[leftmargin=2em]
\item the symmetry transports faces, interiors, reflections, and feasibility
  conditions correctly;
\item every raw case belongs to a symmetry orbit with a chosen representative;
\item the obstruction theorem for the representative transports to the raw
  case.
\end{enumerate}
Thus symmetry never replaces exhaustive coverage; it only reduces the amount of
generated evidence needed to obtain it.

The public generated interface is semantic.  Rather than requiring public
modules to expose all raw certificate data, generated coverage should expose
theorems whose conclusions are case impossibility statements.  Interval
coverage packages these theorems over half-open rank ranges \([l,h)\).  The
module \texttt{Generated/ExhaustiveCoverage.lean} contains adapters from such
interval theorems to the global coverage structures used by the conditional
main theorem.

The translation side now has a particularly compact semantic target.  For a
rank interval \([l,h)\), the theorem
\[
\texttt{AllTranslationGoodCoverageOnRange}(l,h)
\]
asserts that every rank in the interval whose pair word has identity linear
part, and every sign mask satisfying GoodDirection, is eliminated.  The
general GoodDirection theorem then handles the remaining masks: if a feasible
translation orbit existed, it would necessarily be a GoodDirection mask.  In
this way the generated translation theorem is allowed to focus only on the
survivors of a general geometric positivity test.

The descriptor-language adapters give several equivalent ways to prove this
same interval theorem.  A generated file may prove a descriptor coverage
statement directly from semantic GoodDirection hypotheses, or it may prove the
Boolean version using \texttt{goodDirectionAtRankBool}.  Lean-proved adapters
then convert either version to the same public interval theorem.  This
flexibility is useful because some generated shards, meaning small generated
files covering separate pieces of the finite range, are naturally expressed as
Boolean classifiers, while the final mathematical theorem is stated in terms
of ordinary propositions.

\section{Trusted Boundary and Generated Evidence}

The trusted part of the proof consists of:
\begin{itemize}[leftmargin=2em]
\item the exact coordinate model of the cuboctahedron;
\item the reflection and unfolding theorems;
\item the rank/unrank enumeration of valid pair words;
\item soundness theorems for non-identity certificates;
\item soundness theorems for translation, Farkas, source-Farkas, and
  two-source certificates;
\item GoodDirection necessity for feasible translation orbits;
\item the equivalence between GoodDirection and its finite Boolean checker;
\item Walsh sign-polynomial and symbolic denominator soundness theorems;
\item descriptor-family adapters from source rows and row templates to
  semantic translation coverage;
\item generated coverage theorems whose types state that every finite case is
  killed.
\end{itemize}

External Python and legacy C++ programs are outside the trusted boundary.  They
may discover patterns, emit candidate certificates, or produce diagnostics.
They may be wrong.  Lean must check every certificate or family theorem that is
used in the final proof.

At the present snapshot, the active public generated surface is deliberately
small.  \texttt{Generated/AllGenerated.lean} imports the computable-classifier
bridge and a public evidence marker.  A marker is a small Lean module recording
which generated-evidence root is currently exposed to the main project.  This
avoids older all-in-one generated imports that caused memory pressure.  The
current marker records a bounded interval,
\[
[0,8),
\]
whose proof-carrying root is outside the default Lake source tree at
\[
\texttt{evidence/public\_interval\_shards/Shard000000000\_000000008/VerifiedRoot.lean}.
\]
This demonstrates the public interval shape but is not the final exhaustive
coverage theorem over all \(97{,}297{,}200\) pair-word ranks.

Subsequent checked test modules have strengthened the translation-family
interface.  In particular, the file
\[
\texttt{SourceIndexStateDescriptorLanguage.lean}
\]
now proves adapters between semantic and Boolean GoodDirection coverage,
between descriptor coverage and source-row coverage, and from descriptor
coverage to \texttt{AllTranslationGoodCoverageOnRange}.  A generated classifier
test module packages \(125\) descriptor families selected from the rank window
\([0,5000)\), and the focused guarded build of that test completed within an
8 GiB memory guard.  These tests are not global coverage, but they show that
the current public theorem surface can be fed by compact descriptor families
rather than by replaying one rank-mask proof at a time.

\section{Formal Statement}

The final theorem is expected to have the shape
\[
\texttt{\textasciitilde{} exists o : BilliardOrbit14,}
\]
\[
\texttt{\quad o.Nonsingular /\ o.Periodic /\ o.TouchesEachFaceExactlyOnce.}
\]
The current public theorem is the conditional version:
\[
\texttt{conditional\_cuboctahedron\_no\_omnihedral}
  : \texttt{ExhaustiveGeneratedCoverage}\to \text{the above conclusion}.
\]
There are also branch-level conditional theorems:
\begin{align*}
&\texttt{conditional\_generated\_nonidentity\_complete},\\
&\texttt{conditional\_generated\_translation\_complete},\\
&\texttt{conditional\_no\_unfolded\_omni\_starting\_xp}.
\end{align*}
The newer semantic route proves analogous results from
\texttt{SemanticExhaustiveGeneratedCoverage}.  On the translation branch, the
principal semantic input is interval coverage of GoodDirection survivors,
\[
\texttt{AllTranslationGoodCoverageOnRange}(l,h),
\]
which can now be obtained from descriptor-family coverage in either the
semantic or Boolean GoodDirection form.

\section{Reproducibility}

For the current conditional proof surface, the expected verification commands
are:
\begin{verbatim}
lake build
grep -R "sorry\|admit\|axiom\|native_decide\|unsafe" Cuboctahedron || true
lake env lean Cuboctahedron/PrintConditionalAxioms.lean
\end{verbatim}
The final unconditional proof should replace the last line by
\begin{verbatim}
lean Cuboctahedron/PrintAxioms.lean
\end{verbatim}
once the final theorem and full generated coverage root are present.

\section{Statement on AI Use}

The use of AI formal-methods agents was central to this project.  The author
developed the mathematical plan and directed the proof-engineering process.
AI agents were used to build Lean scaffolding, implement exact certificate
checkers, generate and refine proof infrastructure, search for compression
strategies, and maintain the large generated-evidence pipeline.  The trusted
mathematical artifact remains the Lean-checked proof: external conversations,
scripts, diagnostics, and generated suggestions are not trusted unless their
claims are checked by Lean.

\section{Acknowledgments}

The author thanks the Lean and \mathlib communities for the theorem prover and
library infrastructure that make this project possible.  Further
acknowledgments for discussions, hardware, and tool support will be added in a
later revision.

\appendix

\section{Lean Names}

\begin{center}
\begin{tabular}{ll}
\toprule
Paper concept & Lean name \\
\midrule
Period-14 billiard orbit & \texttt{BilliardOrbit14} \\
Face sequence & \texttt{Step14 -> Face} \\
Omnihedral sequence & \texttt{IsOmniSeq} \\
Started sequence at \(X_+\) & \texttt{StartsXp} \\
Unfolded feasibility & \texttt{UnfoldedFeasible} \\
Pair word & \texttt{PairWord} \\
Valid pair word & \texttt{ValidPairWord} \\
Pair-word unranking & \texttt{unrankPairWord} \\
Number of pair words & \texttt{numPairWords} \\
Translation sign mask & \texttt{SignMask} \\
GoodDirection predicate & \texttt{GoodDirectionAtRank} \\
GoodDirection Boolean checker & \texttt{goodDirectionAtRankBool} \\
Strict linear row & \texttt{StrictLin2} \\
Farkas certificate & \texttt{FarkasCert} \\
Farkas checker & \texttt{checkFarkas} \\
Two-source support & \texttt{TwoSourceFarkasSupport} \\
Descriptor family & \texttt{SourceIndexStateFamilyDescriptor} \\
Descriptor coverage & \texttt{SourceIndexStateDescriptorGoodCoverageOnRange} \\
Boolean descriptor coverage & \texttt{SourceIndexStateDescriptorBoolCoverageOnRange} \\
Translation GoodDirection interval coverage & \texttt{AllTranslationGoodCoverageOnRange} \\
Walsh subcube & \texttt{MaskSubcube} \\
Walsh quadratic record & \texttt{WalshQuadratic} \\
Conditional main theorem & \texttt{conditional\_cuboctahedron\_no\_omnihedral} \\
\bottomrule
\end{tabular}
\end{center}

\section{Raw Search Counts}

The following counts are useful as orientation.  The proof does not trust them
unless they are reflected in Lean-checked coverage.

\begin{center}
\begin{tabular}{lr}
\toprule
Quantity & Count \\
\midrule
Valid pair words & \(97{,}297{,}200\) \\
Non-identity pair words & \(94{,}829{,}112\) \\
Identity-linear pair words & \(2{,}468{,}088\) \\
Translation sign assignments & \(157{,}957{,}632\) \\
Valid omnihedral orbits found & \(0\) \\
\bottomrule
\end{tabular}
\end{center}

\section{Generated Evidence Roadmap}

The final generated evidence theorem should provide, at minimum:
\begin{enumerate}[leftmargin=2em]
\item non-identity killed coverage over all ranks with \(M\ne I\);
\item translation killed coverage over all ranks and sign masks with \(M=I\);
\item a GoodDirection adapter showing that generated translation evidence only
  needs to cover masks satisfying the GoodDirection necessary condition;
\item descriptor or source-row family coverage proving
  \texttt{AllTranslationGoodCoverageOnRange} over the required intervals;
\item interval composition over \([0,\texttt{numPairWords})\);
\item optional symmetry transport, with Lean-checked orbit coverage, wherever
  symmetry is used for compression.
\end{enumerate}
The preferred public target is semantic coverage: generated roots should expose
small theorems of impossibility over intervals and families, not large public
tables of raw certificates.

\section{Current Diagnostic Checkpoints}

The following checkpoints are useful for understanding the present state of
the generated-evidence pipeline.  They are not part of the theorem unless their
claims are reflected in Lean-checked coverage imported by the final root.

\begin{center}
\begin{tabular}{p{0.58\linewidth}r}
\toprule
Checkpoint & Value \\
\midrule
GoodDirection survivors in ranks \([0,100000)\) covered by the diagnostic
row-relation census & \(39{,}710\) \\
Calibration-window GoodDirection survivors covered after the expanded
row-template catalog & \(63{,}725\) \\
Representative symbolic row-family survivors & \(4{,}779\) \\
Representative symbolic row families & \(126\) \\
Descriptor classifier families selected from ranks \([0,5000)\) & \(125\) \\
\bottomrule
\end{tabular}
\end{center}

The last checkpoint is the most recent proof-interface test.  It demonstrates
that descriptor families can be used to feed
\texttt{AllTranslationGoodCoverageOnRange} through Lean-checked adapters,
including the Boolean GoodDirection route, without exposing the underlying
rank-mask arithmetic as the public theorem interface.

\begin{thebibliography}{9}

\bibitem{tabachnikov}
S. Tabachnikov,
\emph{Geometry and Billiards},
Student Mathematical Library, Vol. 30,
American Mathematical Society, 2005.

\bibitem{masur}
H. Masur,
Closed trajectories for quadratic differentials with an application to
billiards,
\emph{Duke Mathematical Journal} 53 (1986), no. 2, 307--314.

\bibitem{boshernitzan}
M. Boshernitzan, G. Galperin, T. Kruger, and S. Troubetzkoy,
Some remarks on periodic billiard orbits in rational polygons,
\emph{Commentarii Mathematici Helvetici} 73 (1998), 177--194.

\bibitem{bedaride-rao}
N. Bedaride and M. Rao,
Regular simplices and periodic billiard orbits,
\emph{Journal of Modern Dynamics} 6 (2012), no. 2, 253--265.

\bibitem{cekic-georgiev-mukherjee}
M. Cekic, B. Georgiev, and M. Mukherjee,
Polyhedral billiards, eigenfunction concentration and almost periodic control,
\emph{Communications in Mathematical Physics} 383 (2021), 377--405.

\bibitem{lean4}
L. de Moura and S. Ullrich,
The Lean 4 theorem prover and programming language,
in \emph{Automated Deduction -- CADE 28},
Lecture Notes in Computer Science 12699, Springer, 2021, pp. 625--635.

\bibitem{mathlib}
The mathlib Community,
The Lean mathematical library,
in \emph{Proceedings of the 9th ACM SIGPLAN International Conference on
Certified Programs and Proofs}, 2020, pp. 367--381.

\end{thebibliography}

\end{document}
